### 2018-10-14 article.tex
Draft #1 - 1

\documentclass{article}

\usepackage[backend=biber, style=numeric,autocite=plain,sorting=none]{biblatex}
\addbibresource{bibliography.bib}

\usepackage{amsmath}

\title{Platform for Anomaly Detection in Time-Series}

\begin{document}

\maketitle

\section*{Abstract}

Monitoring complex cyber-physical systes generate large amounts of data in the form of time series. Unfortunately, collected data may contain erroneous values caused by various events (e.g. device defects, communication errors, environmental changes, etc), which may effect the stability and controllability of the system. Therefore, a variety of methods were developed in order to identify and eliminate erroneous data.

The goal of this paper is to present a platform that integrates a number of functionalities necessary in the process of anomaly detection, from data gathering, preprocessing towards various anomaly detection techniques and visualization methods.The purpose of this tool is to allow a developer to test, select and fine tune different algorithms that best fit anomaly detection in a given domain.

Anomaly detection in time series uses various methods such as basic ones inspired from signal processing, methods that are based on time, space or functional correlations, methos based on system theory, or on pattern recognition. These methods perform differently on different data sets and for differrent types of anomalies. There isn't a ``universal recipe'' that works in all cases. Data analysts must choose the right approach for anomaly detection, taking into consideration the characteristics of data, the characteristics of the monitored system and the type of anomaly. Our platform tries to cover most of the detection techniques used in time series and aims at providing data analysts with an environment that allows comparisons between the results of different anomaly detection methods.

To demonstrate the utility of the platform, we present a series of experiments done with different methods for anomaly detection on time series and evaluate their results.

\section{Introduction}

We rely on computer and automation systems to manage complex tasks. They are used in many fields and have many applications. Computer systems are used in industrial, academic, commercial and financial applications because of their speed and reliability. Because a defect in such a system can lead to large monetary loss and potentially even loss of life, we need systems that monitor other systems. These systems need to be able to detect faults and anomalous behavior in some way. Algorithms that detect anomalous behavior are therefore necessary and important.

Other applications of anomaly detecting systems include systems that monitor our health. We would like to identify as soon as possible the onset of any disease. By monitoring our health we may be able to spot individuals that have a high risk of contracting some disease and allow medical professionals to act in time to improve their quality of life.

Most anomaly detection algorithms were developed for a given problem or a given range of problems. While creating general algorithms is really hard and may not even be possible, it is best to try many different approaches. The problem we are trying to solve is the lack of a designated platform or tool that can aid in deciding which anomaly detection algorithm works best for a particular problem.

In this paper we propose a platform that enables a user to test a variety of anomaly detection algorithms, with emphasis on time series data. This data has the form $X = \{x_t \in {\rm I\!R} : \forall t \ge 0\}$. Furthermore, we will use the following functional definition of what an anomaly is: An anomaly is a data point or set of data, which is significantly different from all other data points or sets. This means that in order to define an anomaly, one must first have a notion of nominal data. The term anomaly is relative and can not be applied to a datapoint independently of any dataset.

The following chapters are structured as follows: In Section \ref{relwork} we describe previous and related work on the subject. Next, in Section \ref{methods} we describe the types of anomaly detection methods provided by our platform. We will test these methods in Section \ref{experiments}. This will be followed by a discussion on the future research directions and implications in Section \ref{discussion}.

\section{Related Work \label{relwork}}

Most methods for anomaly detection are developed for a given field or have a specific application. In \cite{nnfd1994} and \cite{MMAD2006} the authors developed methods to detect anomalies in airplane data. In \cite{ghad2018} similar methods were developed for IoT.

In \cite{outlierSurvey2014} a host of algorithms and techniques are described and categorized. The authors found that while some algorithms in different fields are very similar, most algorithms are hard to generalize. They also note, that numerous formulations of anomaly detection problems are not suffieciently explored i.e. it is not known how well some algorithms perform in a field that was not intended for that algorithm.

We believe that there is a real advantage in a platform that offers a variety of anomaly detection methods to the user. One can test the performance of a number of anomaly detection algoritmhs for some given test data. By having as many implementations of these algorithms as possible, the user can easily test as many algorithms as she wants with minimal effort and cost.

Efforts to bundle up different anomaly detection algorithms have already begun. In \cite{egads2015} the authors introduced an open source, generic framework for detecting anomalies in large scale time-series data.

In \cite{tpad2018} we propose a platform, that offers tools for data visualization, filtering and classification for a variety of data formats, including but not limited to time series data.

\section{Methods \label{methods}}

In this section we will present some anomaly detection methods.

\subsection{Outlier Detection Methods}

The first type of problem is concerned with classifying each element as anomalous or nominal. We could define a function $label$ that labels a datapoint $x \in X$ either as anomalous or nominal:
$$
    label: X \times x \rightarrow \{Nominal, Anomaly\}
$$

If it is possible to create a predicate function $p: x \rightarrow \{Nominal, Anomaly\}$ that can accurately detect anomalies, the implementation of the problem becomes trivial: $label(X, x) = p(x)$. Nontheless, such methods can be used to detect obvious anomalies such as extreme temperature levels or very high blood pressure.

If such a predicate function is not possible, or doesn't meet the correctness requirements, a more complex approach can be used. Since the time series is generated by a generative process, one could hope to be able to accurately describe the underlying process, and create a model of the system:
$$
    f: t \rightarrow x^*
$$

Given such a model, one can label the anomalies based on some treshold given a distance metric $d: x \times x \rightarrow {\rm I\!R}$. If the distance between the predicted value $x^*$ and the actual value $x$ is greater than some treshold $d_{max}$, $x$ can be considered an anomaly:
$$
    p(x_t) =
        \begin{cases}
            \text{Anomaly} & \text{if } d(x, f(t)) \ge d_{max} \\
            \text{Nominal} & \text{otherwise}
        \end{cases}
$$

In practice, this is almost never possible. Instead the model will predict the new values based on the old observed values: $f : x^n \rightarrow x^*, n \ge 1$. If $n < |X|$, we use a sliding window approach, where we generate a prediction for the next value based on the actual old values. This is useful if we can model the time series using an auto-regressive model.

\subsection{Change Point Detection}

\subsection{Artificial Intelligence Classification}

\subsection{Anomalous Time Series Detection}

\section{Experiments \label{experiments}}

\section{Discussion \label{discussion}}

\section*{Bibliography}

\printbibliography

\end{document}

### 2018-10-21 article.tex
Added images and corrected spelling

@@
 
-Monitoring complex cyber-physical systes generate large amounts of data in the form of time series.
+Monitoring complex cyber-physical systems generate large amounts of data in the form of time series.
 Unfortunately, collected data may contain erroneous values caused by various events (e.g.
@@
 
-Anomaly detection in time series uses various methods such as basic ones inspired from signal processing, methods that are based on time, space or functional correlations, methos based on system theory, or on pattern recognition.
-These methods perform differently on different data sets and for differrent types of anomalies.
+Anomaly detection in time series uses various methods such as basic ones inspired from signal processing, methods that are based on time, space or functional correlations, methods based on system theory, or on pattern recognition.
+These methods perform differently on different data sets and for different types of anomalies.
 There isn't a ``universal recipe'' that works in all cases.
@@
 This means that in order to define an anomaly, one must first have a notion of nominal data.
-The term anomaly is relative and can not be applied to a datapoint independently of any dataset.
+The term anomaly is relative and can not be applied to a data-point independently of any dataset.
 
@@
 In \cite{ghad2018} similar methods were developed for IoT.
+In \cite{ADDSCIL} methods for detecting anomalies in Big Data are presented.
+In \cite{NDTSDII1996} methods are used to detect anomalies in industrial machineries.
 
@@
 The authors found that while some algorithms in different fields are very similar, most algorithms are hard to generalize.
-They also note, that numerous formulations of anomaly detection problems are not suffieciently explored i.e.
+They also note, that numerous formulations of anomaly detection problems are not sufficiently explored, i.e.
 it is not known how well some algorithms perform in a field that was not intended for that algorithm.
@@
 We believe that there is a real advantage in a platform that offers a variety of anomaly detection methods to the user.
-One can test the performance of a number of anomaly detection algoritmhs for some given test data.
+One can test the performance of a number of anomaly detection algorithms for some given test data.
 By having as many implementations of these algorithms as possible, the user can easily test as many algorithms as she wants with minimal effort and cost.
@@
 
-\section{Methods \label{methods}}
+\section{Time series anomaly detection in the platform \label{methods}}
 
-In this section we will present some anomaly detection methods.
+In this section we will present some anomaly detection methods implemented in the platform. We will give a brief description for each one accompanied by an example.
 
@@
 
-The first type of problem is concerned with classifying each element as anomalous or nominal. We could define a function $label$ that labels a datapoint $x \in X$ either as anomalous or nominal:
+The first type of problem is concerned with classifying each element as anomalous or nominal. We could define a function $label$ that labels a data-point $x \in X$ either as anomalous or nominal:
 $$
@@
 If it is possible to create a predicate function $p: x \rightarrow \{Nominal, Anomaly\}$ that can accurately detect anomalies, the implementation of the problem becomes trivial: $label(X, x) = p(x)$.
-Nontheless, such methods can be used to detect obvious anomalies such as extreme temperature levels or very high blood pressure.
+Nonetheless, such methods can be used to detect obvious anomalies such as extreme temperature levels or very high blood pressure.
 
@@
 
-Given such a model, one can label the anomalies based on some treshold given a distance metric $d: x \times x \rightarrow {\rm I\!R}$.
-If the distance between the predicted value $x^*$ and the actual value $x$ is greater than some treshold $d_{max}$, $x$ can be considered an anomaly:
+Given such a model, one can label the anomalies based on some threshold given a distance metric $d: x \times x \rightarrow {\rm I\!R}$.
+If the distance between the predicted value $x^*$ and the actual value $x$ is greater than some threshold $d_{max}$, $x$ can be considered an anomaly:
 $$

### 2018-10-21 article.tex
Changed style to better fit the requirements

@@
-\documentclass{article}
+\documentclass[a4paper, 10pt]{article}
+
+\usepackage{times}
+
+\usepackage{graphicx}
+\graphicspath{{images/}}
 
@@
 
-\title{Platform for Anomaly Detection in Time-Series}
+\title{\huge Platform for Anomaly Detection in Time-Series}
 
@@
 
-\section*{Abstract}
+\noindent \hrulefill
 
-Monitoring complex cyber-physical systems generate large amounts of data in the form of time series.
-Unfortunately, collected data may contain erroneous values caused by various events (e.g.
-device defects, communication errors, environmental changes, etc), which may effect the stability and controllability of the system.
-Therefore, a variety of methods were developed in order to identify and eliminate erroneous data.
+\small
+\textit{
+Abstract: The goal of this paper is to present a platform that integrates a number of functionalities necessary in the process of anomaly detection, from data gathering, preprocessing towards various anomaly detection techniques and visualization methods.The purpose of this tool is to allow a developer to test, select and fine tune different algorithms that best fit anomaly detection in a given domain.
+To demonstrate the utility of the platform, we present a series of experiments done with different methods for anomaly detection on time series and evaluate their results.
+}
 
-The goal of this paper is to present a platform that integrates a number of functionalities necessary in the process of anomaly detection, from data gathering, preprocessing towards various anomaly detection techniques and visualization methods.The purpose of this tool is to allow a developer to test, select and fine tune different algorithms that best fit anomaly detection in a given domain.
+\noindent \hrulefill
 
-Anomaly detection in time series uses various methods such as basic ones inspired from signal processing, methods that are based on time, space or functional correlations, methods based on system theory, or on pattern recognition.
-These methods perform differently on different data sets and for different types of anomalies.
-There isn't a ``universal recipe'' that works in all cases.
-Data analysts must choose the right approach for anomaly detection, taking into consideration the characteristics of data, the characteristics of the monitored system and the type of anomaly.
-Our platform tries to cover most of the detection techniques used in time series and aims at providing data analysts with an environment that allows comparisons between the results of different anomaly detection methods.
+\textit{
+Keywords: Anomaly Detection
+}
 
-To demonstrate the utility of the platform, we present a series of experiments done with different methods for anomaly detection on time series and evaluate their results.
+\normalsize
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Outlier}
+    \label{fig:outlier}
+    \caption{This is a horse.}
+\end{figure}
+
 \subsection{Change Point Detection}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{ChangePoint}
+    \label{fig:changepoint}
+    \caption{This is a raisin.}
+\end{figure}
 

### 2018-10-21 article.tex
Added space after comma.

@@
 \textit{
-Abstract: The goal of this paper is to present a platform that integrates a number of functionalities necessary in the process of anomaly detection, from data gathering, preprocessing towards various anomaly detection techniques and visualization methods.The purpose of this tool is to allow a developer to test, select and fine tune different algorithms that best fit anomaly detection in a given domain.
+Abstract: The goal of this paper is to present a platform that integrates a number of functionalities necessary in the process of anomaly detection, from data gathering, preprocessing towards various anomaly detection techniques and visualization methods.
+The purpose of this tool is to allow a developer to test, select and fine tune different algorithms that best fit anomaly detection in a given domain.
 To demonstrate the utility of the platform, we present a series of experiments done with different methods for anomaly detection on time series and evaluate their results.

### 2018-10-22 article.tex
Added Workflow picture and other modifications.

@@
 \textit{
-Abstract: The goal of this paper is to present a platform that integrates a number of functionalities necessary in the process of anomaly detection, from data gathering, preprocessing towards various anomaly detection techniques and visualization methods.
+Abstract: The goal of this paper is to present a platform that integrates a number of functionalities necessary in the process of anomaly detection, from preprocessing towards various anomaly detection techniques and visualization methods.
 The purpose of this tool is to allow a developer to test, select and fine tune different algorithms that best fit anomaly detection in a given domain.
-To demonstrate the utility of the platform, we present a series of experiments done with different methods for anomaly detection on time series and evaluate their results.
+To demonstrate the utility of the platform, we present a series of experiments done with different methods for anomaly detection on time-series and evaluate their results.
 }
@@
 \textit{
-Keywords: Anomaly Detection
+Keywords: Anomaly Detection, time-series
 }
@@
 The following chapters are structured as follows: In Section \ref{relwork} we describe previous and related work on the subject.
-Next, in Section \ref{methods} we describe the types of anomaly detection methods provided by our platform.
+Next, in Section \ref{platform} we give a brief description of the developed platform and its features.
+In Section \ref{methods} we describe the types of anomaly detection methods provided by our platform.
 We will test these methods in Section \ref{experiments}.
@@
 In \cite{ADDSCIL} methods for detecting anomalies in Big Data are presented.
-In \cite{NDTSDII1996} methods are used to detect anomalies in industrial machineries.
+In \cite{NDTSDII1996} methods are used to detect anomalies in industrial machinery.
 
@@
 
+Efforts to bundle up different anomaly detection algorithms have already begun.
+In \cite{egads2015} the authors introduced an open source, generic framework for detecting anomalies in large scale time-series data.
+
+In \cite{tpad2018} a platform is proposed, that offers tools for data visualization, filtering and classification for a variety of data formats, including but not limited to time series data.
+
 We believe that there is a real advantage in a platform that offers a variety of anomaly detection methods to the user.
@@
 
-Efforts to bundle up different anomaly detection algorithms have already begun.
-In \cite{egads2015} the authors introduced an open source, generic framework for detecting anomalies in large scale time-series data.
+\section{The Platform \label{platform}}
 
-In \cite{tpad2018} we propose a platform, that offers tools for data visualization, filtering and classification for a variety of data formats, including but not limited to time series data.
+In this section we present the anomaly detection platform and it's functionalities.
 
-\section{Time series anomaly detection in the platform \label{methods}}
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{Workflow}
+\label{fig:workflow}
+\caption{Workflow of the thing.}
+\end{figure}
 
-In this section we will present some anomaly detection methods implemented in the platform. We will give a brief description for each one accompanied by an example.
+\section{Anomaly Detection Techniques \label{methods}}
+
+In this section we give short descriptions of the anomaly detection methods present in the platform.
 
@@
 The first type of problem is concerned with classifying each element as anomalous or nominal. We could define a function $label$ that labels a data-point $x \in X$ either as anomalous or nominal:
-$$
-    label: X \times x \rightarrow \{Nominal, Anomaly\}
-$$
+$$c_x = label(x,X)$$
+$$c_x \in \{Nominal, Anomaly\}$$
 
-If it is possible to create a predicate function $p: x \rightarrow \{Nominal, Anomaly\}$ that can accurately detect anomalies, the implementation of the problem becomes trivial: $label(X, x) = p(x)$.
-Nonetheless, such methods can be used to detect obvious anomalies such as extreme temperature levels or very high blood pressure.
+One might be able to label if an element is an anomaly by setting global lower and upper bounds.
+This can be used to detect obvious anomalies such as extreme temperature levels or very high blood pressure.
 
@@
 Since the time series is generated by a generative process, one could hope to be able to accurately describe the underlying process, and create a model of the system:
-$$
-    f: t \rightarrow x^*
-$$
+$$x_t^* = f(t)$$
 
-Given such a model, one can label the anomalies based on some threshold given a distance metric $d: x \times x \rightarrow {\rm I\!R}$.
-If the distance between the predicted value $x^*$ and the actual value $x$ is greater than some threshold $d_{max}$, $x$ can be considered an anomaly:
+Given such a model, one can label the anomalies based on some threshold given a dstance metric $d: X \times X \rightarrow {\rm I\!R}$.
+If the distance between the predicted value $x_t^*$ and the actual value $x_t$ is greater than some threshold $d_{max}$, $x_t$ can be considered an anomaly:
 $$
@@
         \begin{cases}
-            \text{Anomaly} & \text{if } d(x, f(t)) \ge d_{max} \\
+            \text{Anomaly} & \text{if } d(x_t, f(t)) \ge d_{max} \\
             \text{Nominal} & \text{otherwise}
@@
 $$
+
+An ilustration of this can be seen in Figure \ref{fig:outlier}.
 
@@
 \label{fig:outlier}
-    \caption{This is a horse.}
+\caption{The upper graphs shows measured values from a process, and the lower figure is the classification.
+While the signal stays withing a certain range of the model, represented here by the upper and lower bounds, the signal is considered nominal.
+Points that are outside this ``band of normality'' are considered anomalies.}
 \end{figure}
@@
     \label{fig:changepoint}
-    \caption{This is a raisin.}
+    \caption{In the top most graph, we can see the observed process.
+In this case it is a pure sine wave.
+In the middle graphs we can see the coefficient of the model, which in this case is just a function of it's frequency, since it is enough to perfectly describe the process.
+We can consider as anomaly either the change point, either all the points where the model is outside some bounds.}
 \end{figure}

### 2018-10-22 article.tex
Second Draft Ready.

@@
 
+\usepackage{cite}
+
 \usepackage{graphicx}
 \graphicspath{{images/}}
-
-\usepackage[backend=biber, style=numeric,autocite=plain,sorting=none]{biblatex}
-\addbibresource{bibliography.bib}
 
@@
 
-In this section we present the anomaly detection platform and it's functionalities.
+In this section we present the anomaly detection platform and its functionalities for time-series data.
+
+In Figure \ref{fig:workflow} the envisioned workflow is presented.
+The user starts by loading in some data that she wishes to analyze.
+She can now choose to visualize the data using the visualization tools provided by the platform.
+The user can apply transformations to the dataset either to smooth out the data or in an attempt to improve the classification results.
+Next, a user can try out a host of anomaly detection algorithms.
+Each one will have a set parameters.
+These can be fine tuned to improve the classification quality.
+This can be measured using classification metrics, such as precision and recall.
 
@@
 \includegraphics[width=\textwidth]{Workflow}
+\caption{Workflow of the classification process.}
 \label{fig:workflow}
-\caption{Workflow of the thing.}
 \end{figure}
@@
 If such a predicate function is not possible, or doesn't meet the correctness requirements, a more complex approach can be used.
-Since the time series is generated by a generative process, one could hope to be able to accurately describe the underlying process, and create a model of the system:
+Since the time-series is generated by a generative process, one could hope to be able to accurately describe the underlying process, and create a model of the system:
 $$x_t^* = f(t)$$
+where $x_t^*$ is the predicted value of $x$ at time $t$. \\
 
@@
 
-An ilustration of this can be seen in Figure \ref{fig:outlier}.
+An ilustration of this can be seen in Figure \ref{fig:outlier}. \\
 
-In practice, this is almost never possible.
-Instead the model will predict the new values based on the old observed values: $f : x^n \rightarrow x^*, n \ge 1$.
-If $n < |X|$, we use a sliding window approach, where we generate a prediction for the next value based on the actual old values.
+Other regressive models, that use the past values to predict the new values are also used: $x_t^* = f(x_{t-1}, x_{t-2}, \dots, x_{t-n}), n \ge 1$.
+If given that the data starts at $t=0$ and $t-n > 0$, we use a sliding window approach, where we generate a prediction for the next value based on the actual old values.
 This is useful if we can model the time series using an auto-regressive model.
@@
 \includegraphics[width=\textwidth]{Outlier}
-\label{fig:outlier}
 \caption{The upper graphs shows measured values from a process, and the lower figure is the classification.
@@
 Points that are outside this ``band of normality'' are considered anomalies.}
+\label{fig:outlier}
 \end{figure}
@@
 
+Change Point detection focuses on the underlying model of the process.
+The data is generated by a generative process $f(t, p)$, where $p \in {\rm I\!R}^n$ are the parameters of the model.
+That process is considered to be the nominal generative process.
+There is also an error $e(t)$ associated with the process.
+The error function is usually considered to be white noise.
+$$x_t = f(t, p) - e(t)$$
+
+This detection methods focuses on more long term changes compared to regular outlier detection methods.
+When a change in the system behavior is observed, it is considered an anomalous behavior.
+In other words, we constantly update the parameters of the process $p_n$, and compare it with the previous parameters $p_{n-1}$.
+$p_n$ is considered an anomalous behavior if $|p_n - p_{n-1}| > d_p$, where $d_p \in {\rm I\!R}$ is a threshold value.
+
+An example of this is illustrated in Figure \ref{fig:changepoint}.
+
 \begin{figure}
@@
 \includegraphics[width=\textwidth]{ChangePoint}   
-    \label{fig:changepoint}
 \caption{In the top most graph, we can see the observed process.
@@
 In the middle graphs we can see the coefficient of the model, which in this case is just a function of it's frequency, since it is enough to perfectly describe the process.
-We can consider as anomaly either the change point, either all the points where the model is outside some bounds.}
+We consider as anomaly either the change point, either all the points where the model is outside some bounds.}
+\label{fig:changepoint}
 \end{figure}
@@
 
+In this section we will perform some experiments on some data, but we have not yet decided what we are going to do.
+
 \section{Discussion \label{discussion}}
 
-\section*{Bibliography}
+As a conclusion, we think it is a good idea to pull together all the anomaly detection methods used for time-series into a single platform.
 
-\printbibliography
+\bibliography{refs} 
+\bibliographystyle{ieeetr}
 

### 2018-10-27 article.tex
Working on draft 3

@@
 \centering
-\includegraphics[width=\textwidth]{Workflow}
+\includegraphics[width=0.7\textwidth]{Workflow}
 \caption{Workflow of the classification process.}
@@
 
-In this section we give short descriptions of the anomaly detection methods present in the platform.
+In this section we give short descriptions of some of the possible anomaly types. For each anomaly type we present some methods for detecting such an anomaly.
 
-\subsection{Outlier Detection Methods}
+\subsection{Outliers}
 
@@
 
-One might be able to label if an element is an anomaly by setting global lower and upper bounds.
-This can be used to detect obvious anomalies such as extreme temperature levels or very high blood pressure.
+\subsubsection{Bounded Method}
+
+One might be able to label elements as anomalies by setting global lower and upper bounds. This can be used to detect obvious anomalies such as extreme temperature levels or very high blood pressure.
+
+\subsubsection{Model Distance Method}
 
@@
 
-Other regressive models, that use the past values to predict the new values are also used: $x_t^* = f(x_{t-1}, x_{t-2}, \dots, x_{t-n}), n \ge 1$.
-If given that the data starts at $t=0$ and $t-n > 0$, we use a sliding window approach, where we generate a prediction for the next value based on the actual old values.
-This is useful if we can model the time series using an auto-regressive model.
-
 \begin{figure}
 \centering
-\includegraphics[width=\textwidth]{Outlier}
+\includegraphics[width=0.7\textwidth]{Outlier}
 \caption{The upper graphs shows measured values from a process, and the lower figure is the classification.
@@
 \end{figure}
+
+\subsubsection{Linear Approximation Method}
+
+\subsubsection{Autoregressive Method}
+
+Other regressive models, that use the past values to predict the new values are also used: $x_t^* = f(x_{t-1}, x_{t-2}, \dots, x_{t-n}), n \ge 1$.
+If given that the data starts at $t=0$ and $t-n > 0$, we use a sliding window approach, where we generate a prediction for the next value based on the actual old values.
+This is useful if we can model the time series using an auto-regressive model.
 
@@
 \centering
-\includegraphics[width=\textwidth]{ChangePoint}   
+\includegraphics[width=0.7\textwidth]{ChangePoint}   
 \caption{In the top most graph, we can see the observed process.

### 2018-10-29 article.tex
I don't know what changed
Something did.

@@
 
-The first type of problem is concerned with classifying each element as anomalous or nominal. We could define a function $label$ that labels a data-point $x \in X$ either as anomalous or nominal:
+The first type of problem is concerned with classifying each element as anomalous or nominal. We define a function $label$ that labels a data-point $x \in X$ either as anomalous or nominal:
 $$c_x = label(x,X)$$
 $$c_x \in \{Nominal, Anomaly\}$$
+
+There are online and offline versions of each method.
+Offline methods usually perform better, since for each data point we can use both past and future values.
+These methods may be applied in batch, or can be used online, but the detection will have a delay.
 

### 2018-10-29 article.tex
Draft 3

@@
 \usepackage{cite}
+
+\usepackage{placeins}
 
@@
 
-In this section we present the anomaly detection platform and its functionalities for time-series data.
+In this section we present the anomaly detection platform and its functionalities for time-series data. In Figure \ref{fig:workflow} the envisioned workflow is presented.
 
-In Figure \ref{fig:workflow} the envisioned workflow is presented.
 The user starts by loading in some data that she wishes to analyze.
-She can now choose to visualize the data using the visualization tools provided by the platform.
-The user can apply transformations to the dataset either to smooth out the data or in an attempt to improve the classification results.
+Because data can be stored in many different formats, the platform was designed to work with excel, Comma Separated Values (CSV) and Attribute-Relation File Format (ARFF) files, alongside MAT files.
+
+The user can choose to visualize the data using the visualization tools provided by the platform.
+We provide tools both for time-series and non time-series data.
+For time-series data tools for plotting, histograms and Fourier decomposition can be used.
+
+The user can apply transformations to the dataset either to smooth out the data in an attempt to improve the classification results.
+For this we provide high-pass, low-pass and band-pass filters.
+We call this step preprocessing.
+
 Next, a user can try out a host of anomaly detection algorithms.
-Each one will have a set parameters.
+Each one will have a set of parameters.
 These can be fine tuned to improve the classification quality.
-This can be measured using classification metrics, such as precision and recall.
+The quality of the classification can be measured using classification metrics.
+The confusion matrix can be generated as well as some derived metrics such as precision and recall.
+
+When the user is satisfied with the classification, she can use the parameters of the classification method either for future tests or to implement a specialized system for monitoring anomalies.
 
@@
 \end{figure}
+
+\FloatBarrier
 
@@
 
+In the platform we provide a method that classifies a point as an anomaly if it is outside certain bounds.
+This can be done even with any derivative of the function.
+We call this \emph{Bounded Derivative Method}.
+BDM is defined as:
+$$
+	BDM(x) = 
+	\begin{cases}
+       \text{Nominal} & \text{if } B^+ > f(x)^{(n)} > B^-\\
+    	\text{Anomaly} & \text{otherwise}
+    \end{cases}
+$$
+
+where $B^+$ and $B^-$ are the upper and lower bounds, and $f(x)^{(n)}$ is the $n$-th derivative of the signal. An example of the usage can be seen in Figure \ref{fig:BDM}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{BDM}
+\caption{A signal with a higher than average disturbance in the middle is detected by the method.}
+\label{fig:BDM}
+\end{figure}
+
+\FloatBarrier
+
+Even though this detection is trivial, it can be used to detect many types of anomalies. This is because of the fact that anomalies appear outside normal working conditions.
+
 \subsubsection{Model Distance Method}
@@
 
-An ilustration of this can be seen in Figure \ref{fig:outlier}. \\
+Although such a model is highly useful, sometimes in practice it is good enough to consider only the neighboring points.
+For illustration we will use a method inspired from \cite{MMAD2006}.
+This method is known as the \emph{Double Sided Median Method} for anomaly detection.
+By using a sliding window, we calculate the mean of the values inside the window, and if a given value is outside the allowed bounds, it is considered an anomaly.
+
+$$
+	DSMM(x_t) = 
+		\begin{cases}
+            \text{Anomaly} & \text{if } |f(x_t) - mean(f(x_{t-k}), \dots, f(x_{t+k}))| > d \\
+            \text{Nominal} & \text{otherwise}
+        \end{cases}
+$$
+
+An example can be seen in Figure \ref{fig:DSMM}. \\
 
@@
 \centering
-\includegraphics[width=0.7\textwidth]{Outlier}
-\caption{The upper graphs shows measured values from a process, and the lower figure is the classification.
-While the signal stays withing a certain range of the model, represented here by the upper and lower bounds, the signal is considered nominal.
-Points that are outside this ``band of normality'' are considered anomalies.}
-\label{fig:outlier}
+\includegraphics[width=0.49\textwidth]{DSMM}
+\caption{In the top figure, we can see the raw signal.
+In green we see the distance of each point from the mean of the sliding window.
+The black line from the middle plot is the distance limit.
+If the values falls outside the maximum distance, that point is considered to be an anomaly, as can be seen in the bottom plot.}
+\label{fig:DSMM}
 \end{figure}
 
+\FloatBarrier
+
 \subsubsection{Linear Approximation Method}
+
+TODO
 
@@
 
-\subsection{Artificial Intelligence Classification}
-
 \subsection{Anomalous Time Series Detection}

### 2018-11-03 article.tex
Working on draft 3
Added some text for anomaly detection.

@@
 
+\subsection{Data Loading}
+
 The user starts by loading in some data that she wishes to analyze.
@@
 
+\subsection{Plotting}
+
 The user can choose to visualize the data using the visualization tools provided by the platform.
 We provide tools both for time-series and non time-series data.
-For time-series data tools for plotting, histograms and Fourier decomposition can be used.
+For time-series data we provide tools for plotting, histograms and Fourier decomposition.
+
+\subsection{Preprocessing}
 
@@
 We call this step preprocessing.
+
+\subsection{Anomaly Detection}
 
@@
 When the user is satisfied with the classification, she can use the parameters of the classification method either for future tests or to implement a specialized system for monitoring anomalies.
+
+In the next chapter we present the types of anomalies together with the detection methods used to detect them.
 
@@
 
-In this section we give short descriptions of some of the possible anomaly types. For each anomaly type we present some methods for detecting such an anomaly.
+In this section we give short descriptions of some of the possible anomaly types. For each anomaly type we present some methods for detecting such an anomaly. \\
+
+In principle anomaly detection is similar to system identification.
+We would like to have a model of the system working in the ``nominal'' state.
+In the artificial intelligence fields it can be considered as a classification problem, since we assign labels to data.
+
+Anomaly detection is slightly different from these in one regard: anomalies are generally rare, and comprise a very small percent of the data.
+Some detection methods go for the system identification route, and try to find a model that best fits the nominal data, as well as excludes the anomalous points.
+Some methods go for the classification route, where artificial intelligence methodologies are used for classification.
+
+The problem is that the quality of classification provided by the above mentioned methods depend on the size of the dataset.
+The essence of anomaly detection lies in the fact, that anomalies are rare events, and most datapoints in the training data are ``nominal'' points.
+
+Some datasets do not even contain anomalies, and algorithms are expected to learn the nominal functioning of the system from ``clean data''.
+We call a dataset that doesn't contain anomalies a clean dataset.
+\\
+
+Another classification can be whether or not the methods are online or offline methods.
+Offline methods usually perform better, since for each data point we can use both past and future values.
+These methods may be applied in batch, or can be used online, but the detection will have a delay.
+\\
+
+Given these constraints one can start to classify the anomaly types and detection algorithms.
 
@@
 $$c_x \in \{Nominal, Anomaly\}$$
-
-There are online and offline versions of each method.
-Offline methods usually perform better, since for each data point we can use both past and future values.
-These methods may be applied in batch, or can be used online, but the detection will have a delay.
 
@@
 \includegraphics[width=0.7\textwidth]{BDM}
-\caption{A signal with a higher than average disturbance in the middle is detected by the method.}
+\caption{A signal with a higher than average disturbance in the middle is detected by the method. Forward difference is used as an approximation for the derivative.}
 \label{fig:BDM}
@@
 
-If such a predicate function is not possible, or doesn't meet the correctness requirements, a more complex approach can be used.
+If such a predicate function is not possible, a more complex approach can be used.\\
+
 Since the time-series is generated by a generative process, one could hope to be able to accurately describe the underlying process, and create a model of the system:

### 2018-11-05 article.tex
Started doing tests
Man, I am really falling behind on this crap.

@@
 
-\subsection{Anomalous Time Series Detection}
-
 \section{Experiments \label{experiments}}
 
-In this section we will perform some experiments on some data, but we have not yet decided what we are going to do.
+In this section we will compare a number of anomaly detection methods, and compare their results. \\
+
+We will use precision and recall to compare the methodologies presented.
+
+First dataset
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\textwidth]{testplot1}   
+\caption{Description pending.}
+\label{fig:test1}
+\end{figure}
+
+\FloatBarrier
+
+\begin{center}
+\begin{tabular}{| c | c | c |}
+\hline
+Method                      & Precision & Recall \\ \hline
+Bounded Derrivative (d = 0) & 1         & 0.3678 \\ \hline
+Bounded Derrivative (d = 1) & 0.6       & 0.0689 \\ \hline
+Median Method               & 0.33      & 0.0919 \\ \hline
+\end{tabular}
+\end{center}
 
